Changes in Patterns of Geometric Sign Groupings as a Model of Social Dynamics Across Aurignacian Phases

### 2025-03-21 endofquarterpaper.qmd
replace , with ; in citation set in intro

The Upper Paleolithic period experienced a rise in more complex social organizations across Europe [@Boyd_Silk_2020] coupled with biological evolution. Another distinctive quality of the Upper Paleolithic is the large rise in artwork, both parietal and mobile. Archaeologists have frequently relied upon that artwork as a proxy for cultural difference in order to understand the shifting patterns of social organization and diversity of the Upper Paleolithic [@baker2024evidence, @Sauvet_2018, @Kuhn_Stiner_2007, @Vanhaeren_d]. Considered to be the starting point of the Upper Paleolithic, the Aurignacian technocomplex (ca. 43 - 30 k BP) is commonly linked to the dispersal of Homo sapiens across Western Eurasia, and is characterized by lithic and osseus markers [@Tejero_Grimaldi_2015; @Tartar_2012], some of which are decorated with geometric signs [@dutkiewicz2020]. In this paper, we investigate the social organization dynamics by analyzing the distribution and variation of geometric sign types found on mobile objects from the Upper Paleolithic in Europe across four phases of the Aurignacian period.

@@

## Cultural Group Detection
```{r fig.height= 9, fig.width= 13}
#| label: fig-seriation
#| fig-cap: "Seriation analysis of the data, with group membership indicated by color. Restricted group is in yellow, broader group is in blue."

>>>>

The Upper Paleolithic period experienced a rise in more complex social organizations across Europe [@Boyd_Silk_2020] coupled with biological evolution. Another distinctive quality of the Upper Paleolithic is the large rise in artwork, both parietal and mobile. Archaeologists have frequently relied upon that artwork as a proxy for cultural difference in order to understand the shifting patterns of social organization and diversity of the Upper Paleolithic [@baker2024evidence; @Sauvet_2018; @Kuhn_Stiner_2007; @Vanhaeren_d]. Considered to be the starting point of the Upper Paleolithic, the Aurignacian technocomplex (ca. 43 - 30 k BP) is commonly linked to the dispersal of Homo sapiens across Western Eurasia, and is characterized by lithic and osseus markers [@Tejero_Grimaldi_2015; @Tartar_2012], some of which are decorated with geometric signs [@dutkiewicz2020]. In this paper, we investigate the social organization dynamics by analyzing the distribution and variation of geometric sign types found on mobile objects from the Upper Paleolithic in Europe across four phases of the Aurignacian period.

@@

## Cultural Group Detection

```{r fig.height= 9, fig.width= 13}
#| label: fig-seriation
#| fig-cap: "Seriation analysis of the data, with group membership indicated by color. Restricted group is in yellow, broader group is in blue."
